---
title: "STAT 512 Final Project Team 11"
author: "Nick Gunady"
date: "12/4/2021"
output: html_document
knit: (function(input_file, encoding) {
  out_dir <- 'docs';
  rmarkdown::render(input_file,
 encoding=encoding,
 output_file=file.path(dirname(input_file), out_dir, 'index.html'))})
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
rm(list = ls())
filedir <- getwd()
```

```{r packages}
library(magrittr)
library(tinytex)
library(ALSM)
library(stringr)
library(stats)

#plotly Regression packages
library(reshape2) # to load tips data
library(tidymodels) # for the fit() function
library(plotly)
data(tips)
```

CA demographics dataset can be downloaded from: https://www.countyhealthrankings.org/app/california/2021/downloads

```{r data wrangling}
#main CA demographics dataset
demo <- read.csv(file="demographics.csv")

 #main working df
df.demo <- demo[2:58,c(3,4,7,8,9,10,11,14,17,18,19,20,21,22,23)]
df.demo[,c(2,4,6,8,12)] <- df.demo[,c(2,4,6,8,12)] / 100 #change % columns to decimal

#main covid case data from jhu.edu
covid <- read.csv(file="CA-2020-cov-data.csv") %>% dplyr::rename(Cases = Case.Count.2020)

df.main <- left_join(df.demo, covid, by="County") %>% dplyr::select(Cases,County,-X,everything())
df.main <- df.main[,1:16]
```

```{r initial plotting}
plot(df.main)
```
Experiment 1: Full MLR
``` {r full MLR}
Y <- df.main$Cases -> Confirmed.COVID19.Cases
X1 <- df.main$NUM.FOOD.INSECURE -> Food.Insecurity
X2 <- df.main$PERCENT.ADULT.DIABETES -> Percent.Adult.Diabetes
X3 <- df.main$PERCENT.OBESITY -> Percent.Adult.Obesity
X4 <- df.main$PERCENT.LIMITED.ACCESS -> Access.to.Healthy.Food
X5 <- df.main$MEDIAN.HOUSEHOLD.INCOME -> Median.Household.Income
X6 <- df.main$AVG.DAILY.PM2.5 -> Air.Pollution

df.full <- cbind(Y,X1,X2,X3,X4,X5,X6) %>% as.data.frame

mod.full <- lm(Confirmed.COVID19.Cases~Food.Insecurity+Percent.Adult.Diabetes+Percent.Adult.Obesity+Access.to.Healthy.Food+Median.Household.Income+Air.Pollution, df.full)

summary(mod.full)
anova(mod.full)
Anova
```

From the ANOVA results of the full MLR, it is clear further model refinement is required. Therefore, we plot the MLR model criteria with plotmodel.s function, part of the ALSM package. Additionally, the Best Subsets Algorithm (also from ALSM) is utilized to determine the best model for this experiment. 
```{r full MLR model selection}
#Plot Criterias for Model Selection
plotmodel.s(df.full[,2:7],df.full$Y)
```

```{r MLR BestSub}

#Best Subsets Algorithm for Model Selection
bs.full <- BestSub(df.full[2:7],df.full$Y,num=1) %>% as.data.frame
bs.full

#Which r^2 and r^2 adjusted are greatest?
r2.max <- bs.full$r2 %>% max
nrow(bs.full==r2.max)

r2adj.max <- bs.full$r2.adj %>% max
nrow(bs.full==r2adj.max)

Cp.min <- bs.full$Cp %>% min
nrow(bs.full==Cp.min)

AIC.min <- bs.full$AICp %>% min
nrow(bs.full==AIC.min)

SBC.min <- bs.full$SBCp %>% min
nrow(bs.full==SBC.min)

PRESSp.min <- bs.full$PRESSp %>% min
nrow(bs.full==PRESSp.min)
```


Experiment 2: Food insecurity as a factor for COVID-19 cases
H0: b1 = 0
Ha: b1 ~= 0
```{r food insecurity}
Y <- df.main$Cases
X <- df.main$NUM.FOOD.INSECURE

df.fi <- cbind(X,Y) %>% as.data.frame

mod.fi <- lm(Y~X, df.fi)

summary(mod.fi)
anova(mod.fi)
```

```{r food insecurity plots}
#plot linear model data
plot(mod.fi)

#plot data with regression line
plot(X,Y,pch = 16, cex = 1.3, col = "blue", main = "Confirmed COVID-19 Cases vs. Food Insecurity", xlab = "Food Insecurity", ylab = "Confirmed COVID-19 Cases")
abline(mod.fi)

#Plotly
fig.fi <- plot_ly(df.fi, x = ~X, y = ~Y, 
               type = 'scatter',
               alpha=0.65,
               mode='markers',
               name='COVID-19 Cases') %>% 
  add_lines(x=~X,y=fitted(mod.fi),name='Regression Line') %>%
  layout(title='COVID-19 Cases vs. Food Insecurity',
         xaxis = list(title='Number of Food Insecure Adults'),
         yaxis = list(title='Confirmed COVID-19 Cases'))
fig.fi
```

Experiment 3: Comparing Food Insecurity vs. Health factors such as obesity and diabetes.
H0: b1 = 0, b2 = 0, b3 = 0
Ha: not H0

``` {r food insecurity health}
Y <- df.main$Cases -> Confirmed.COVID19.Cases
X1 <- df.main$NUM.FOOD.INSECURE -> Food.Insecurity
X2 <- df.main$PERCENT.ADULT.DIABETES -> Percent.Adult.Diabetes
X3 <- df.main$PERCENT.OBESITY -> Percent.Adult.Obesity
X4 <- df.main$PERCENT.LIMITED.ACCESS -> Access.to.Healthy.Food
X5 <- df.main$MEDIAN.HOUSEHOLD.INCOME -> Median.Household.Income
X6 -> df.main$AVG.DAILY.PM2.5 -> Air.Pollution

df.fih <- cbind(Y,X1,X2,X3,X4,X5,X6) %>% as.data.frame

mod.fih <- lm(Confirmed.COVID19.Cases~Food.Insecurity+Percent.Adult.Diabetes+Percent.Adult.Obesity, df.fih)

summary(mod.fih)
anova(mod.fih)
```
```{r Experiment 2 Plots}
plot(mod.fih)

#Plotly
fig.fih <- plot_ly(df.fih, x = ~X1+X2+X3, y = ~Y, 
               type='scatter', 
               alpha=0.65, 
               mode='markers',
               name='COVID-19 Cases') %>%
  add_lines(x=~X1+X2+X3,y=fitted(mod.fih),name='Regression Line') %>%
  layout(title='COVID-19 Cases vs. Food Insecurity+Obesity+Diabetes',
         xaxis = list(title='Food Insecurity + Obesity + Diabetes'),
         yaxis = list(title='Confirmed COVID-19 Cases'))
fig.fih
```

Experiment 3: Unemployment as a factor for COVID-19 cases
H0: b1 = 0
Ha: b1 ~= 0
```{r unemployment}
Y <- df.main$Cases
X <- df.main$NUM.UNEMPLOYED

df <- cbind(X,Y) %>% as.data.frame

mod.unemp <- lm(Y~X, df)

summary(mod.unemp)
anova(mod.unemp)
```

``` {r unemployment plots}
#plot linear model data
plot(mod.unemp)

#Plotly
fig.unemp <- plot_ly(df, x = ~X, y = ~Y, 
               type='scatter', 
               alpha=0.65, 
               mode='markers',
               name='COVID-19 Cases') %>%
  add_lines(x=~X,y=fitted(mod.unemp),name='Regression Line') %>%
  layout(title='COVID-19 Cases vs. Unemployment',
         xaxis = list(title='Number of Unemployed Adults'),
         yaxis = list(title='Confirmed COVID-19 Cases'))
fig.unemp
```
